The Openscapes Flywheel: A framework for managers to facilitate and scale inclusive Open science practices

### 2022-08-19 manuscript.qmd
first pass at refs, through engage. cc @erinmr

---
title: "The Openscapes Flywheel: A framework for managers to facilitate and scale inclusive Open science practices"
date: now
date-format: "DD MMM YYYY"
toc: true
author:
  - name: Erin M. Robinson
    orcid: 0000-0001-9998-0114
    email: erinmr@metadatagamechangers.com
    affiliations:
      - name: Metadata Game Changers
        city: Boulder
        state: CO
  - name: Julia S. Stewart Lowndes
    orcid: 0000-0003-1682-3872
    affiliations:
      - name: National Center for Ecological Analysis & Synthesis, University of California
        city: Santa Barbara
        state: CA
csl: nature.csl
bibliography: references.bib
---

@@

Solutions to large-scale environmental and social challenges require radical collaboration that blends technology and people. The Open science movement is answering that call by transforming how we work together. While often the focus is on researchers, software developers, and data, code, and publications, the transition to Open involves investment across agencies and organizations of all kinds. One critical and often-overlooked part of transitioning to Open is that it requires support from managers and leadership, who do not need to be Open science experts but who can imagine the return on investment this new way of working provides, and support their teams through time and resources to learn new skills, take risks, and even fail through the process. In this paper we describe our work developing the Openscapes Flywheel with NASA Earth science data centers, using the flywheel concept where transformations occur from relentlessly pushing a giant, heavy flywheel that builds momentum at every turn. We offer three practical ways managers can support their teams and grow morale and technical capacity across their organizations: (1) Engage bright spots; (2) Empower a learning culture; (3) Amplify Open science leaders. Together these lessons create kinder and more inclusive organizations that are more resilient as data volumes increase and software evolve, and dispel the hero myth of science. Engaging, empowering and amplifying efforts to support people while modernizing our workforce is the way that we will meet the world\'s greatest challenges, from climate change to social justice to democracy. We can already see how working this way increases inclusion, reproducibility, and resilience within and across government, academic, and non-profit organizations, and are hopeful for what we can do together in growing this movement.

@@

##  Transition Towards Open

Solutions to large-scale environmental and social challenges require radical collaboration, but have been limited by the hero model, where science is competitive, siloed, and rewards an individual hero, often a white man. While we know that science has never been done by individuals, Science perpetuated the hero myth for hundreds of years (Elkins-Tanton, 2021). Open science is a global movement that offers transparency, efficiency, reproducibility, reuse, and community with intentional focus on inclusion - the vision of a new paradigm for science  ([UNESCO](https://unesdoc.unesco.org/ark:/48223/pf0000379949.locale=en)). The Open science movement was codified by the Budapest Open Access Initiative (2001), and after decades of grassroots community building work and technology development, we are seeing the turn toward Open science with the recent publication of the UNESCO Recommendations on Open Science ([UNESCO](https://unesdoc.unesco.org/ark:/48223/pf0000379949.locale=en)), increasing numbers of funders coordinated by the Open Research Funder Group (ORFG), Open access publishing and now large government institutions like NASA\'s Transform to Open Science (NASA, 2021). While Open science is not new, the transition to Open practices in a digital, data intensive world, requires investment in a new way of working (Royal Society, 2012).

Meanwhile, at many scientific organizations, managers and leadership are facing a need to modernize -  upskill the workforce, update technology, foster a more inclusive culture, support a distributed workforce, create and share reports more efficiently, transparently, and reproducibly, and steward and analyze more data, to name a few - often without a roadmap and with flat budgets. Open science offers consistency and resilience for managers, yet we are still seeing many research teams continuing to do science as usual, because they don\'t have the bandwidth to transition (Plomp 2022). Researchers feel squeezed from all sides, responsible to do more with less, while gaps widen within teams and organizations about who \"is technical\", or interacts with data, code, and project management. Largely, if researchers transition to Open science it is often self-taught, done on the side, as a volunteer effort and often not rewarded in traditional academic or government systems.

Transitioning towards Open science isn\'t just something to feel good about and it isn\'t all or nothing. It doesn\'t happen by going to a single workshop and it isn\'t a box to check. It is about people as well as technology - enabling people to collaborate and interoperate with technology to achieve their science goals.

@@

Openscapes works with professional research teams in government, academia, and nonprofits, and we help these teams modernize their workflows and shift culture within their teams and organizations by incorporating Open science into their existing work - the job to be done. We consider the transition to Open science as a journey through a landscape full of data, software, and communities, with many existing pathways forward (Figure 1). We help people transition from feeling like the sad animals in the lower left foreground, who can feel stuck and lonely with this expansive landscape of possibility just out of view, to safely exploring and navigating this landscape safely together, with reduced bushwacking and trampling. Openscapes welcomes research groups at the trailhead, helping them see themselves as a team and identify where they are all starting from. There are different trails and environments that teams work in and while there are many common pieces and skills to reuse and remix, each adventure will be different. \

![](https://lh4.googleusercontent.com/9iRNDqKazjThZ4EU_xiF7n2uS9Mq494X515NQ82dNRX6NsaT9D_uOjHG7NkA9zH5b0JG3CG_XO9sVKb8_u0lXddZXBhtEarNJK9-MzRXLov2_8lctbsGNAqdveZ6Et9SQWZpcL78en4Dd3Chyvc9jg)

Figure 1: Open science landscape and pathways. Longer description: existing paths and creating new paths together. Art by Dr. Allison Horst\

Openscapes Champions, our flagship program, mentors cohorts of research teams in their transition to Open and kinder science (Openscapes, 2022). We have led over 16 cohorts with over 112 teams (\> 520 individuals) over 2-4 months as they identify their trailhead and define their pathway together (Lowndes et al. 2019). Meeting remotely for 1.5 hours twice monthly, we provide dedicated learning time for researchers to explore collaborative approaches for project management, data analysis, and stewardship, and practice incorporating them into their existing work while also bonding and reflecting with colleagues within their team and cohort. Grounded in meeting researchers where they are (Lowndes et al. 2017), Openscapes Champions helps teams tackle big challenges as smaller steps. It combines Open science and open source software with deliberate approaches for inclusion and kindness through facilitation, coaching, teaching, and community building, so that folks are empowered by daily decisions and habits for their work and collaborations that boost morale, while also being more efficient. \

With this relatively small amount of support, we have seen Champions teams take this investment and make great strides on their pathway to Open science (Beck et al. 2019, Bastille et al. 2021). And long before this shows up in scientific publications, it shows up in team morale and workflows: preparing government reports and saving hundreds of would-be emails and would-be file versions and email attachments (Rios); overhauling graduate courses through sharing artwork as open educational resources reused by thousands globally (Horst); thoughtfully having difficult conversations for succession planning so that decades of knowledge and data are not lost as folks retire (Holmes); and reusing role-modeled facilitation and teaching approaches to help teams transition to remote and hybrid work and support each other during the Pandemic (Fay). These are a few examples over a one year period - it is amazing what people can do when they are supported.\

Throughout our work, we've heard from researchers across career stages and paths that finding their team and feeling less lonely and stuck has been one of the most important parts of Openscapes Champions - away from the hero model. And just like how Open science doesn\'t depend on heroes, Openscapes does not either. We have developed a framework that when repeated and reinforced using the flywheel concept (Collins 2011), helps individuals become networked Open science leaders who support others to break down silos, make lasting, meaningful, and transformative changes within and across their organizations and amplified far beyond. This work means slowing down to speed up, which we are seeing return dividends.

@@

NASA Earth Science Data Systems missions collect Earth data, including sea ice, physical oceanography, vegetation and many other parameters that are used by researchers around the world for many purposes, including pressing questions in ecology and environmental science that are part of addressing the climate crisis. Currently, NASA is migrating this data to the cloud. NASA\'s science data centers (Distributed Active Archive Centers, or DAACs) are supporting researchers using Earthdata to migrate their workflows to the Cloud. These are paths that have not been created before but build from existing ones.

Through a NASA grant to Openscapes, we are supporting and strengthening the community of NASA DAAC staff that are already creating cloud learning resources - a project that blends technology and people for radical collaboration. We help this community of cross-DAAC staff - the NASA Openscapes Mentors - learn existing paths and develop new paths together by identifying common priorities, needs, and gaps to fill, while testing cloud learning resources and investing in teaching. Our project increases visibility of Mentors and research teams with cloud and strengthens the community of practice around open science at NASA. In doing so, with the Mentors we are piloting and iteratively co-developing the NASA Openscapes Framework.

@@

**Welcome bright spots.** As we designed our project, we intentionally avoided coming in from the outside to change the culture. Instead, we focused on identifying and connecting the bright spots (ref) - people across the organization already modeling the desired change you want to see. We reached out to managers with a welcoming description of our project and asked them to share the opportunity with people in their group who were working with research teams moving to the Cloud and who were interested in teaching and collaborating. We then provided a short self-nomination form so that folks had agency to opt-in, and asked them to sign up as pairs and trios so they would have a team from their DAAC to join the larger cohort. Opting-in was a critical first investment from both managers and staff. These nominees became the Mentor Cohort, representing five DAACs in year one and now seven in year two.

@@

We provided scaffolding and facilitation for regular meetings, and managed software where Mentors could collaborate. Each DAAC had a different way of working, but Mentors were all agreeable to trying to use common places. Our shared software began with Google Drive and Slack. All software we used included time spent to learn how to use it, and how we wanted to use it together. We also prioritized sharing stories via talks, blogs, and Twitter, so that this articulation work (ref) of community building was visible and valued, and also didn\'t land on the Mentors.

Part of creating space and place is to remove or reduce barriers to participation. This took two forms with the Mentors - managerial and technical. On the managerial side, we were clear from the beginning that NASA Openscapes would require \~8 hours/month, and that that investment would align with their jobs. We advocated to their managers that this is not extra, unpaid labor, but it is part of people\'s jobs and they are evaluated on these contributions (see Amplifying section for more). We also confirmed from NASA leadership that everyone could use GitHub as part of NASA\'s commitment to Open science, and we created a GitHub organization we managed for the group.

On the technical side, many in our group had not ever accessed NASA Earthdata in the Cloud. The primary block seemed to be who would pay for access, how much will it cost to use the Cloud, and how will researchers pay for cloud computing resources? Those questions are important, but they are not the first questions for staff supporting research teams, so we took them off the table. We partnered with 2i2c (ref) to host a JupyterHub that is supported by AWS credits. The 2i2c JupyterHub greatly reduced \"time to Cloud\" for Mentors and helped them experience it for the first time and build confidence. With management support to participate in this cross-DAAC activity, job alignment, a shared interest in collaborating and learning together, along with an established space, place, and barriers removed, the group could focus on how we wanted to co-develop and teach tutorials together.

@@

We have learned many lessons through our work with the Openscapes Mentors and Champions and generalized and tested them as part of our Openscapes Flywheel (fig 2) that gains momentum as it is repeated with each cycle. This guides our own work as we iterate with teams, and our goal is that others can continue to utilize it as they help their organizations transition to Open science. We have been investing in Mentors from the government, like NASA, NOAA, and local state agencies, as well as academic and non-profit groups who can help others navigate them safely and efficiently. This Flywheel is a concrete approach to help people find their pathways as they move into the Open science landscape, focusing on inclusion and community building at every step of the way. As one Mentor, Ileana Fenwick says, it\'s about being \"the community member who you would want to meet at the beginning of your journey\" ([tweet](https://twitter.com/StefanieButland/status/1552720392556466176)).

@@

1.  **Welcome bright spots** - Initially, you don\'t need to find the most resistant toughest collaborators; Work with the willing to build momentum and help others feel the change you envision together. Bright spots are folks that want to do better, even if they don\'t know what that means yet. There is no purity test or technical yard-stick.

@@

3.  **Invest in learning and trust** - Everyone has something to learn, ask, teach; don\'t need to be an expert in everything. Cultivate psychological safety, growth mindset, reinforcing the power of yet. Slowing down to speed up - taking stock at the trailhead to confirm that everyone is there and ready.

@@

6.  **Inspire others and grow the movement** - it doesn\'t have to be a grand act; share your story or a tool tip that inspires other bright spots and the flywheel turns again.

While this Flywheel cycle is the ideal, in practice it often feels like iterative progress forward and back. However, our experience with the Openscapes Flywheel is that there is an important place for managers to enable the transition towards Open science, and to name these parts involved and see where they have momentum and where they are stuck (\"this is a heavy part of the flywheel\"). Increasingly, researchers and mentors we work with are becoming managers, and managers participating with their teams in the Champions program want to sustain this beyond a single program. We hope the Flywheel can be as useful for others as it has been for us; it guides us on many scales, from our hour-long meetings to our 3-year programs.

## Conclusion

The Open science trail systems, as in real life, don\'t just happen overnight. They become worn in through use or they are intentionally built with intentional investment, and both these grassroots and top-down approaches are true for the long-standing and ongoing efforts of the Open movement. Increasing the number of researchers and teams who can travel these paths safely requires investment in not only the paths themselves but in the people who travel and maintain these paths. This is about all of us acting together as a \"we\". In the book, \'All We Can Save\', Dr. Ayana Elizabeth Johnson and Dr. Katherine Wilkinson write about the meaning of each word in the title for tackling the climate crisis together: \

> \"We\" speaks to the collective, to collaboration, to community, to the relational work at hand.... \"We\" speaks to justice, to how we do the work that needs doing and whose contributions are valued. We cannot, we must not, go it alone.\"

With the Openscapes Approach and Flywheel, together we can create kinder and more inclusive organizations and they dispel the hero myth of science. In addition to bright spot researchers and mentors, there are bright spot managers out there that we are grateful to work with and learn from. These managers see the vision and potential for Open science and have empowered the bright spot mentors and research teams, invested in changing the path and are cultivating a teaching culture to build more resilient organizations. Engaging, empowering and amplifying efforts to support people while modernizing our workforce is the way that we will meet the world\'s greatest challenges, from climate change to social justice to democracy. We can already see how working this way increases the diversity within our organization and beyond and are hopeful for what we can do together in growing this movement.

@@

Julia Stewart Lowndes, PhD is a marine ecologist working at the intersection of actionable environmental science, data science, and open science. She is a Mozilla Fellow and Senior Fellow at the National Center for Ecological Analysis and Synthesis (NCEAS) at the University of California Santa Barbara, having earned her PhD from Stanford University in 2012 studying drivers and impacts of Humboldt squid in a changing climate. Lowndes directs Openscapes after launching the Openscapes Champions Program in 2018. \

# Authors\' Contributions

@@

This work was supported by NASA grant #20-TWSC20-2-0003, entitled, \"Openscapes: Enabling a new era of science on the cloud in response to the 2020 NASA ROSES Element E2.\"

@@

>>>>

---
title: "The Openscapes Flywheel: A framework for managers to facilitate and scale inclusive Open science practices"
toc: true
author:
  - name: Erin M. Robinson
    orcid: 0000-0001-9998-0114
    email: erinmr@metadatagamechangers.com
    affiliations:
      - name: Metadata Game Changers
        city: Boulder
        state: CO
  - name: Julia S. Stewart Lowndes
    orcid: 0000-0003-1682-3872
    affiliations:
      - name: National Center for Ecological Analysis & Synthesis, University of California
        city: Santa Barbara
        state: CA
csl: nature.csl
bibliography: references.bib
---

@@

Solutions to large-scale environmental and social challenges require radical collaboration that blends technology and people. The Open science movement is answering that call by transforming how we work together. While often the focus is on researchers, software developers, and data, code, and publications, the transition to Open involves investment across agencies and organizations of all kinds. One critical and often-overlooked part of transitioning to Open is that it requires support from managers and leadership, who do not need to be Open science experts but who can imagine the return on investment this new way of working provides, and support their teams through time and resources to learn new skills, take risks, and even fail through the process. In this paper we describe our work developing the Openscapes Flywheel with NASA Earth science data centers, using the flywheel concept where transformations occur from relentlessly pushing a giant, heavy flywheel that builds momentum at every turn. We offer three practical ways managers can support their teams and grow morale and technical capacity across their organizations: (1) Engage bright spots; (2) Empower a learning culture; (3) Amplify Open science leaders. Together these lessons create kinder and more inclusive organizations that are more resilient as data volumes increase and software evolve, and dispel the hero myth of science. Engaging, empowering and amplifying efforts to support people while modernizing our workforce is the way that we will meet the world's greatest challenges, from climate change to social justice to democracy. We can already see how working this way increases inclusion, reproducibility, and resilience within and across government, academic, and non-profit organizations, and are hopeful for what we can do together in growing this movement.

@@

## Introduction: Transitioning Towards Open

Solutions to large-scale environmental and social challenges require radical collaboration, but have been limited by the hero model, where science is competitive, siloed, and rewards an individual hero, often a white man. While we know that science has never been done by individuals, Science perpetuated the hero myth for hundreds of years @elkins-tanton2021. Open science is a global movement that offers transparency, efficiency, reproducibility, reuse, and community with intentional focus on inclusion - the vision of a new paradigm for science @unesco2021. The Open science movement was codified by the Budapest Open Access Initiative @budapestopenaccessinitiative2001, and after decades of grassroots community building work and technology development, we are seeing the turn toward Open science with the recent publication of the UNESCO Recommendations on Open Science @unesco2021, increasing numbers of funders coordinated by the Open Research Funder Group @openres, Open access publishing (REF) and now large government institutions like NASA's Transform to Open Science @nasa2022. While Open science is not new, the transition to Open practices in a digital, data intensive world, requires investment in a new way of working @nelson2022 (REF Royal Society, 2012).

Meanwhile, at many scientific organizations, managers and leadership are facing a need to modernize - upskill the workforce, update technology, foster a more inclusive culture, support a distributed workforce, create and share reports more efficiently, transparently, and reproducibly, and steward and analyze more data, to name a few - often without a roadmap and with flat budgets. Open science offers consistency and resilience for managers, yet we are still seeing many research teams continuing to do science as usual, because they don't have the bandwidth to transition @plomp2022. Researchers feel squeezed from all sides, responsible to do more with less, while gaps widen within teams and organizations about who "is technical", or interacts with data, code, and project management. Largely, if researchers transition to Open science it is often self-taught, done on the side, as a volunteer effort and often not rewarded in traditional academic or government systems.

Transitioning towards Open science isn't just something to feel good about and it isn't all or nothing. It doesn't happen by going to a single workshop and it isn't a box to check. It is about people as well as technology - enabling people to collaborate and interoperate with technology to achieve their science goals.

@@

Openscapes works with professional research teams in government, academia, and nonprofits, and we help these teams modernize their workflows and shift culture within their teams and organizations by incorporating Open science into their existing work - the job to be done @openscapesteam. We consider the transition to Open science as a journey through a landscape full of data, software, and communities, with many existing pathways forward (@fig-landscape-horst). We help people transition from feeling like the sad animals in the lower left foreground, who can feel stuck and lonely with this expansive landscape of possibility just out of view, to safely exploring and navigating this landscape safely together, with reduced bushwacking and trampling. Openscapes welcomes research groups at the trailhead, helping them see themselves as a team and identify where they are all starting from. There are different trails and environments that teams work in and while there are many common pieces and skills to reuse and remix, each adventure will be different. \

![Open science as a landscape with many existing pathways to travel safely together intentionally as teams and communities. Artwork by Dr. Allison Horst](images/horst_openscapes_landscape.png){#fig-landscape-horst fig-align="center" width="545"}

\

Openscapes Champions, our flagship program, mentors cohorts of research teams in their transition to Open and kinder science @openscapesteam2022. We have led over 16 cohorts with over 112 teams (\> 520 individuals) over 2-4 months as they identify their trailhead and define their pathway together @lowndes2019. Meeting remotely twice monthly, we provide dedicated learning time for researchers to explore collaborative approaches for project management, data analysis, and stewardship, and practice incorporating them into their existing work while also bonding and reflecting with colleagues within their team and cohort. Grounded in meeting researchers where they are @lowndes2017, Openscapes Champions helps teams tackle big challenges as smaller steps. It combines Open science and open source software with deliberate approaches for inclusion and kindness through facilitation, coaching, teaching, and community building, so that folks are empowered by daily decisions and habits for their work and collaborations that boost morale, while also being more efficient. \

With this relatively small amount of support, we have seen Champions teams take this investment and make great strides on their pathway to Open science @beck2020, @bastille2021. And long before this investment shows up in scientific publications, it shows up in team morale and workflows: preparing government reports and saving hundreds of would-be emails and would-be file versions and email attachments (REF Rios); overhauling graduate courses through sharing artwork as open educational resources reused by thousands globally @horst2022 (REF); thoughtfully having difficult conversations for succession planning so that decades of knowledge and data are not lost as folks retire (REF Holmes); and reusing role-modeled facilitation and teaching approaches to help teams transition to remote and hybrid work and support each other during the Pandemic (REF Fay). These are a few examples over a one year period - it is amazing what people can do when they are supported.\

Throughout our work, we've heard from researchers across career stages and paths that finding their team and feeling less lonely and stuck has been one of the most important parts of Openscapes Champions - away from the hero model. And just like how Open science doesn't depend on heroes, Openscapes does not either. We have developed a framework that when repeated and reinforced using the flywheel concept @collins2011, helps individuals become networked Open science leaders who support others to break down silos, make lasting, meaningful, and transformative changes within and across their organizations and amplified far beyond. This work means slowing down to speed up, which we are seeing return dividends.

@@

NASA Earth Science Data Systems missions collect Earth data, including sea ice, physical oceanography, vegetation and many other parameters that are used by researchers around the world for many purposes, including pressing questions in ecology and environmental science that are part of addressing the climate crisis. Currently, NASA is migrating this data to the cloud. NASA's science data centers (Distributed Active Archive Centers, or DAACs) are supporting researchers using Earthdata to migrate their workflows to the Cloud. These are paths that have not been created before but build from existing ones.

Through a NASA grant to Openscapes, we are supporting and strengthening the community of NASA DAAC staff that are already creating cloud learning resources - a project that blends technology and people for radical collaboration @nasaopenscapesmentors. We help this community of cross-DAAC staff - the NASA Openscapes Mentors - learn existing paths and develop new paths together by identifying common priorities, needs, and gaps to fill, while testing cloud learning resources and investing in teaching. Our project increases visibility of Mentors and research teams with cloud and strengthens the community of practice around open science at NASA. In doing so, with the Mentors we are piloting and iteratively co-developing the NASA Openscapes Framework.

@@

**Welcome bright spots.** As we designed our project, we intentionally avoided coming in from the outside to change the culture. Instead, we focused on identifying and connecting the bright spots @heath2010 - people across the organization already modeling the desired change you want to see. We reached out to managers with a welcoming description of our project and asked them to share the opportunity with people in their group who were working with research teams moving to the Cloud and who were interested in teaching and collaborating. We then provided a short self-nomination form so that folks had agency to opt-in, and asked them to sign up as pairs and trios so they would have a team from their DAAC to join the larger cohort. Opting-in was a critical first investment from both managers and staff. These nominees became the Mentor Cohort, representing five DAACs in year one and now seven in year two.

@@

We provided scaffolding and facilitation for regular meetings, and managed software where Mentors could collaborate. Each DAAC had a different way of working, but Mentors were all agreeable to trying to use common places. Our shared software began with Google Drive and Slack. All software we used included time spent to learn how to use it, and how we wanted to use it together. We also prioritized sharing stories via talks, blogs, and Twitter, so that this articulation work (REF) of community building was visible and valued, and also didn't land on the Mentors.

Part of creating space and place is to remove or reduce barriers to participation. This took two forms with the Mentors - managerial and technical. On the managerial side, we were clear from the beginning that NASA Openscapes would require \~8 hours/month, and that that investment would align with their jobs. We advocated to their managers that this is not extra, unpaid labor, but it is part of people's jobs and they are evaluated on these contributions (see Amplifying section for more). We also confirmed from NASA leadership that everyone could use GitHub as part of NASA's commitment to Open science, and we created a GitHub organization we managed for the group.

On the technical side, many in our group had not ever accessed NASA Earthdata in the Cloud. The primary block seemed to be who would pay for access, how much will it cost to use the Cloud, and how will researchers pay for cloud computing resources? Those questions are important, but they are not the first questions for staff supporting research teams, so we took them off the table. We partnered with 2i2c @holdgraf to host a JupyterHub that is supported by AWS credits. The 2i2c JupyterHub greatly reduced "time to Cloud" for Mentors and helped them experience it for the first time and build confidence. With management support to participate in this cross-DAAC activity, job alignment, a shared interest in collaborating and learning together, along with an established space, place, and barriers removed, the group could focus on how we wanted to co-develop and teach tutorials together.

@@

We have learned many lessons through our work with the Openscapes Mentors and Champions and generalized and tested them as part of our Openscapes Flywheel (fig 2) that gains momentum as it is repeated with each cycle. This guides our own work as we iterate with teams, and our goal is that others can continue to utilize it as they help their organizations transition to Open science. We have been investing in Mentors from the government, like NASA, NOAA, and local state agencies, as well as academic and non-profit groups who can help others navigate them safely and efficiently. This Flywheel is a concrete approach to help people find their pathways as they move into the Open science landscape, focusing on inclusion and community building at every step of the way. As one Mentor, Ileana Fenwick says, it's about being "the community member who you would want to meet at the beginning of your journey" ([tweet](https://twitter.com/StefanieButland/status/1552720392556466176)).

@@

1.  **Welcome bright spots** - Initially, you don't need to find the most resistant toughest collaborators; Work with the willing to build momentum and help others feel the change you envision together. Bright spots are folks that want to do better, even if they don't know what that means yet. There is no purity test or technical yard-stick.

@@

3.  **Invest in learning and trust** - Everyone has something to learn, ask, teach; don't need to be an expert in everything. Cultivate psychological safety, growth mindset, reinforcing the power of yet. Slowing down to speed up - taking stock at the trailhead to confirm that everyone is there and ready.

@@

6.  **Inspire others and grow the movement** - it doesn't have to be a grand act; share your story or a tool tip that inspires other bright spots and the flywheel turns again.

While this Flywheel cycle is the ideal, in practice it often feels like iterative progress forward and back. However, our experience with the Openscapes Flywheel is that there is an important place for managers to enable the transition towards Open science, and to name these parts involved and see where they have momentum and where they are stuck ("this is a heavy part of the flywheel"). Increasingly, researchers and mentors we work with are becoming managers, and managers participating with their teams in the Champions program want to sustain this beyond a single program. We hope the Flywheel can be as useful for others as it has been for us; it guides us on many scales, from our hour-long meetings to our 3-year programs.

## Conclusion: Onward Openly together

The Open science trail systems, as in real life, don't just happen overnight. They become worn in through use or they are intentionally built with intentional investment, and both these grassroots and top-down approaches are true for the long-standing and ongoing efforts of the Open movement. Increasing the number of researchers and teams who can travel these paths safely requires investment in not only the paths themselves but in the people who travel and maintain these paths. This is about all of us acting together as a "we". In the book, 'All We Can Save', Dr. Ayana Elizabeth Johnson and Dr. Katherine Wilkinson write about the meaning of each word in the title for tackling the climate crisis together:\

> "We" speaks to the collective, to collaboration, to community, to the relational work at hand.... "We" speaks to justice, to how we do the work that needs doing and whose contributions are valued. We cannot, we must not, go it alone."

With the Openscapes Approach and Flywheel, together we can create kinder and more inclusive organizations and they dispel the hero myth of science. In addition to bright spot researchers and mentors, there are bright spot managers out there that we are grateful to work with and learn from. These managers see the vision and potential for Open science and have empowered the bright spot mentors and research teams, invested in changing the path and are cultivating a teaching culture to build more resilient organizations. Engaging, empowering and amplifying efforts to support people while modernizing our workforce is the way that we will meet the world's greatest challenges, from climate change to social justice to democracy. We can already see how working this way increases the diversity within our organization and beyond and are hopeful for what we can do together in growing this movement.

@@

Julia Stewart Lowndes, PhD is a marine ecologist working at the intersection of actionable environmental science, data science, and open science. She is a Mozilla Fellow and Senior Fellow at the National Center for Ecological Analysis and Synthesis (NCEAS) at the University of California Santa Barbara, having earned her PhD from Stanford University in 2012 studying drivers and impacts of Humboldt squid in a changing climate. Lowndes directs Openscapes after launching the Openscapes Champions Program in 2018.\

# Authors' Contributions

@@

This work was supported by NASA grant #20-TWSC20-2-0003, entitled, "Openscapes: Enabling a new era of science on the cloud in response to the 2020 NASA ROSES Element E2."

@@

# References
